---
title: "Day5-osmdata"
author: "FG"
date: "11/5/2021"
output: html_document
---

#TidyTuesday week45 & #30DayMapChallenge day5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This time we need to make a map using **OSM which is Open Street Map**, also there is a package {osmdata} that can be used for retrieving data and recursive searching with special functions.


```{r libraries, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(purrr)
library(spData)
library(spDataLarge)

library(osmdata)

library(sf)
library(raster)
```

Let's choose a data set from the thousands available in these two packages:

 - {spData}
 - {spDataLarge}
 
 
The **"census_de"** dataset from {spDataLarge} contains the building blocks for a location analysis. Names of the dataset are in German, but we can easily translate them into English.

```{r}
spDataLarge::census_de%>%
  names
```
 
 
```{r}
census_de%>%count(Einwohner) # population in German
```
 
 
```{r df_names}
my_df<-census_de
names(my_df)<- c("grid_id","x","y",
                 "population","women","mean_age",
                 "under_18","ab65","foreigners",
                 "house_hold_size","vacancy_rate",
                 "residential_bew",
                 "residential_whg")
```



```{r table, message=FALSE, warning=FALSE, paged.print=FALSE}
my_table <- my_df%>% #count(women)
  dplyr::select(population,women,mean_age,house_hold_size)%>%
  mutate_all(list(~ifelse(.%in% c(-1,-9),NA,.)))%>%
  drop_na()%>%
  mutate(class=population,.before = population) %>%
  count(class)%>%
  mutate(population=case_when(class=="1"~"3-250",
                              class=="2"~"250-500",
                              class=="3"~"500-2000",
                              class=="4"~"2000-4000",
                              class=="5"~"4000-8000",
                              class=="6"~">8000"),
         women=case_when(class=="1"~"0-40",
                         class=="2"~"40-47",
                         class=="3"~"47-53",
                         class=="4"~"53-60",
                         class=="5"~">60"),
        mean_age=case_when(class=="1"~"0-40",
                         class=="2"~"40-42",
                         class=="3"~"42-44",
                         class=="4"~"44-47",
                         class=="5"~">47"),
        householdsize=case_when(class=="1"~"1-2",
                         class=="2"~"2-2.5",
                         class=="3"~"2.5-3",
                         class=="4"~"3-3.5",
                         class=="5"~">3.5"))%>%
  count(class,population,women,mean_age,householdsize) %>%
  dplyr::select(-n)

my_table[is.na(my_table)]<-""         
  
my_table_gt <- gt::gt(my_table)  
my_table_gt
```
This data set provides the coordinates (x and y) for representing spatial data, and some other variables such as:

- population
- women
- mean age
- house hold size

which we will be using for making our map. The values of the variables are all **made of dummy variables**. 

The preprocessing part is very important for setting up a dataset which is appropriate for mapping, and making further analysis passing through the EDA (exploratory data analysis) stage up to modeling.


This [source](https://geocompr.robinlovelace.net/location.html#tab:census-desc) provides a case study using this data set and OSM data retrival.

First step is to set the `input` data with assigning `NA` values to the observation that are outside of the boundaries, in our case are (-1 and -9), so that we will have some white spaces.

```{r input_data}
input_tidy <- my_df%>% 
  dplyr::select(x,y,population,women,mean_age,house_hold_size)%>%
  mutate_all(list(~ifelse(.%in% c(-1,-9),NA,.)))

input_tidy%>%drop_na()
```



Now that **the data set is preprocessed** and ready to use we attempt a retrieval of selected features to map with Open Street Map features.


The {osmdata} package is a package for importing OpenStreetMap (OSM) data into R and for making transformation of selected data as required.

Some important functions are: 

- getbb()             Get bounding box
- opq()               Build an Overpass query
- add_osm_feature()   Add a feature
- osmdata_sf()        Return an object in sf format

The `bbox` argument is important as it defines the bounding box with coordinates. It requires 4 points, 2 for latitude and 2 for longitude, as max and min values of the polygon.


An **object of class osmdata**, is an object with OSM components such as:

- points
- lines
- polygons


```{r eval=FALSE, include=FALSE}
osmdata::getbb()
purrr::map()
```
 

In the case of `census_de` data we have x and y as a list of coordinates, we could check their range, or select some city names among those included in the map such as:

- Berlin

This information is different from the "shops" dataset in {spDataLarge}, because is just for "Berlin", it is a **Object of class 'osmdata'**.

Steps:

1. set a desired location
2. check which types of data to add in the map (as lines/circles/points)
3. apply the search

```{r step1}
# 1. 
location <- "Berlin"
```

```{r}
# osmdata::
```



```{r step2, eval=FALSE, include=FALSE}
# 2.
# length(available_features()) # 223
available_features()[60:120]


available_features()[available_features()=="water"]
```

```{r eval=FALSE, include=FALSE}
feature <- "water"
available_tags(feature)

sel_tag<- c("basin","canal","ditch","wastewater","pond","river")
```



```{r eval=FALSE, include=FALSE}
head(available_tags("boundary")) # "amenity"
head(available_tags("pub")) 
```

For example :

    add_osm_feature(key = "...", value = "...")
   
    add_osm_feature(key = "amenity", value = "pub")


Let's try and select "pub" as a key feature, we found 5 points to add in the map, no lines and one polygon.

```{r}
location<- "Berlin"

pub_osm = getbb(location)%>%
            opq() %>%
            add_osm_feature(key = "pub")%>%
            add_osm_feature(key = "water")%>%
            osmdata_sf()
```
It doesn't work for some reasons ("GDAL Error 1: No PROJ.4 translation for source SRS, coordinate transformation initialization has failed.
"). We need to find an alternative, which is **leaflet** !!

```{r geom_sf, eval=FALSE, include=FALSE}
# ggplot() +
#   geom_sf(data = pub_osm$osm_points,
#           inherit.aes = FALSE,
#           color = "midnightblue",
#           size = .4,
#           alpha = .8) 
```

```{r eval=FALSE, include=FALSE}
available_features()[150:223]
available_features()[available_features()=="internet_access"]
available_tags("water")
```


```{r EDA_map}
input_tidy%>%#filter(!is.na(women))%>%
  ggplot()+
  stat_summary_hex(aes(x,y,z=population))
```



Convert xy into coordinates in a geometry
```{r}
my_sf_df <- st_as_sf(input_tidy, coords = c("x", "y"), crs = 31493) 
# CRS (coordinates reference system) used found on https://www.spatialreference.org/ref/?search=Germany&srtext=Search

#input_tidy%>%dim;my_sf_df%>%dim
#coord_xy<- input_tidy[1:2]

#add xy coords
my_sf_df$X <- input_tidy$x
my_sf_df$Y <- input_tidy$y


my_sf_df_short<- my_sf_df[c(1,5:7)]

my_sf_df_short <- my_sf_df_short%>%filter(!is.na(population))

st_coordinates(my_sf_df_short)

```


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
location<- "Berlin"

internet_access_osm = getbb(location)%>%
            opq() %>%
            add_osm_feature(key = "internet_access")

residential_osm = getbb(location)%>%
             opq() %>%
             add_osm_feature(key = "residential")

water_osm = getbb(location)%>%
             opq() %>%
             add_osm_feature(key = "water")


Location_internet_access <- osmdata_sf(internet_access_osm)
# saveRDS(Location_internet_access,"Location_internet_access.rds")
Location_internet_access <- readRDS("Location_internet_access.rds")


Location_residential <- osmdata_sf(residential_osm)
#saveRDS(Location_residential,"Location_residential.rds")
Location_residential <- readRDS("Location_military.rds")

Location_water <- osmdata_sf(water_osm)
#saveRDS(Location_water,"Location_water.rds")
Location_water <- readRDS("Location_water.rds")


# step 04 - extract amenity points/
points_internet_access <- Location_internet_access$osm_points
points_residential<- Location_residential$osm_points
points_water<- Location_water$osm_points

# step 05 - extract from the amenity points, node ID and geometry (lat,  lng)
points_geo_internet_access <- data.frame(Location_internet_access$osm_points$osm_id,
                                         Location_internet_access$osm_points$geometry)

points_geo_residential <- data.frame(Location_residential$osm_points$osm_id,
                                     Location_residential$osm_points$geometry)

points_geo_water <- data.frame(Location_water$osm_points$osm_id,
                              Location_water$osm_points$geometry)

# step 06 - rename first column node ID as osm_id
names(points_geo_internet_access)[1]<- "osm_id"
names(points_geo_residential)[1]<- "osm_id"
names(points_geo_water)[1]<- "osm_id"

# step 07 - merge data frames
points_internet_access <- merge.data.frame(points_internet_access, points_geo_internet_access)
points_residential <- merge.data.frame(points_residential, points_geo_residential)
points_water <- merge.data.frame(points_water, points_geo_water)


# retrieve coordinates in matrix form
coord_internet_access <- data.frame(st_coordinates(points_internet_access$geometry))
points_internet_access$X <- coord_internet_access$X
points_internet_access$Y <- coord_internet_access$Y

coord_residential <- data.frame(st_coordinates(points_residential$geometry))
points_residential$X <- list(coord_residential$X)
points_residential$Y <- list(coord_residential$Y)

coord_water <- data.frame(st_coordinates(points_water$geometry))
points_water$X <- list(coord_water$X)
points_water$Y <- list(coord_water$Y)


library(leaflet)
leaflet() %>%
  addTiles() %>%
  setView(lat=52.517317, lng=13.412364, zoom = 14) %>%
  addProviderTiles(providers$Stamen.Toner)  %>% 
  addProviderTiles(providers$Stamen.TonerLines,
    options = providerTileOptions(opacity = 0.35)) %>%
  addProviderTiles(providers$Stamen.TonerLabels) %>%
  
  # set a box with open street map
  addRectangles(lat1=52.52007,
    lng1=13.38709,
    lng2=13.39707,
    lat2=52.51631,
    fillColor = "transparent") %>%
 addCircleMarkers(data = points_internet_access,
                   lat = ~Y,
                   lng = ~X,
                   radius = 0.5,
                   popup = ~internet_access,
                   color = "yellow3") %>%

  addPolygons(data = points_residential,
                   lat = ~unlist(Y),
                   lng = ~unlist(X),
                   group = geometry,
                   color = "pink4",
                   popup = ~rsidential,
                   fillColor = NA,
                   fillOpacity = 0.2,
              weight = 1.5 ) 
 #>%
 #
 #addPolygons(data = points_water,
 #                 lat = ~unlist(Y),
 #                 lng = ~unlist(X),
 #                 group = geometry,
 #                 color = "midnightblue",
 #                 popup = ~ access,
 #                 fillColor = NA,
 #                 fillOpacity = 0.2,
 #            weight = 1.5
 #            )
```


```{r}
# %>%
#   addLogo(img, url = "https://www.r-project.org/logo/")
```



```{r}
library(devtools)
install_github("wch/webshot")

## load packages
library(htmlwidgets)
library(webshot)


## save html to png
saveWidget(map, "temp.html", selfcontained = FALSE)
webshot("temp.html", file = "Rplot.png", cliprect = "viewport")
```















Or we can set a bbox with the coordinates for a section of the map.

```{r}
#shops_osm = getbb(viewbox=)%>%

       berl_shops<- opq(bbox=c(13.3610,52.5189,13.4154,52.5062)) %>%
        add_osm_feature(key = "shop")%>%
        osmdata_sf()
```


## spDataLarge data ----
Let's say for example that we want to populate our map with shops. We can download the shops dataset, made ready from {spDataLarge} and ? How do we used {osmdata}?
 
let's see how to apply these "shpos" data to a map. This is made of **simple features** and some named variables: 

- osm_id
- shop
- geometry

```{r shops}
spDataLarge::shops
```

Also, other data available from the same package such as: "metro_names".
```{r}
spDataLarge::metro_names
```
 
 
 
 
